---
title: "Data 602 - Class Notes for October 3"
output: html_notebook
editor_options: 
  markdown: 
    wrap: 72
---

Hypothesis Testing about $p$

**Open Question for Example 1:**

$$
{\rm H}_{0}: p = (\leq) 0.40 \hspace{0.5in} {\rm H}_{A}: p > 0.40
$$

Consider the value of $\alpha = 0.0971$ and $beta_{p = 0.45} = 0.7691$.

What would happen if the values of $\alpha$ and $\beta_{p = 0.45}$ if
the null hypothesis were to be rejected in favour of the alternative
hypothesis if $$
\text{Reject} \:\:{\rm H}_{0}: \{X: X \geq 17\} \hspace{0.5in} \text{rather than}\:\: \{X: X \geq 16\}
$$

$$
\begin{eqnarray}
\alpha & = & P(\text{Reject}\:\:{\rm H}_{0} | {\rm H}_{0} \: \text{is true}) \\
       & = & P(X \geq 17 | p = 0.40) \\
       & = & \sum_{x = 17}^{30}{30 \choose x}(0.40)^{x}(0.60)^{30 - x} \\
       & = & 0.0481
\end{eqnarray}
$$

```{r}
sum(dbinom(17:30, 30, 0.40))
```

Similarly,

$$
\begin{eqnarray}
\beta & = & P(\text{Fail to Reject}\:\:{\rm H}_{0} | {\rm H}_{0} \: \text{is false}) \\
       & = & P(X \leq 16 | p = 0.45) \\
       & = & \sum_{x = 0}^{16}{30 \choose x}(0.45)^{x}(0.50)^{30 - x} \\
       & = & 0.8644
\end{eqnarray}
$$

```{r}
sum(dbinom(0:16, 30, 0.45))
```

**Time to Play With Data 1:**

Revisiting part (a) from the practice with stating statistical
hypotheses: $p$ was defined as the proportion of Millennial Canadians
(aged 18 to 34 years of age) who are relying on a financial inheritance
to reach their financial goals. </br> </br> We wish to test that this
proportion equates to **at most half** of the persons in this
group/population. State the statistical hypothesis. The statistical
hypothesis was stated as </br> </br> **Answer:**

1.  We have formulated the statistical hypothesis early in this class:
    $$
    \begin{eqnarray}
    {\rm H}_{0}: p  & \leq (=) & \:0.50 \hspace{0.25in} {\rm \text{(the proportion of Millenials who rely on an inheritance to reach financial goals is at most 50%)}} \\
    {\rm H}_{0}: p & >& \:0.50 \hspace{0.25in} {\rm \text{(the proportion of Millenials who rely on an inheritance to reach financial goals is more than 50%)}} \\
    \end{eqnarray}
    $$ </br> </br>

2 and 3. Collect the data. Although the breakdown of the sample by age
cohort was not provided, let's presume that the sample size is
relatively large and $n_{Millenial} = 759$. From this, the observed
number who indicated that they were relying on an inheritance to achieve
their financial goals is $X_{Obs} = 425$. </br> </br>

4.  The Distribution of the Statistic $X_{Obs}$?

::: pink
```{r}
plot(seq(300, 759, 1), dbinom(seq(300, 759, 1), 759, 0.50), xlab="Values of Sample Sum", ylab="P(X = X_{Obs})", type="h", col="blue", main="Distribution of the Test Statistic")
abline(v = 425, col="red")
```
:::

5.  Computation of the $P$-value. Under the assumption/condition that
    the null hypothesis ${\rm H}_{0}: p = 0.50$ is true, the $P$-value
    is

$$
\begin{eqnarray}
P-\text{value} & = &  P(X > \overbrace{425}^{X_{Obs}} | p = 0.50) \\
               & = & \sum_{x = 426}^{n = 759}{759 \choose x}{0.50^{x}( 1 - 0.50)^{759 - x}}\\
               & = & 0.0004142 \:\: \text{(see R code below)}\\
               & \approx & 0.0004
\end{eqnarray}
$$

::: pink
```{r}
pvalue.ex1 <- 1 - pbinom(425, 759, 0.5) # computes P(X <= 425) with n = 759 and the incorporated assummed value of p = 0.50
pvalue.ex1
```
:::

Note: For a regulated value of $\alpha = 0.05$, the rejection region is

$$
\text{Reject the null hypothesis if} \:\: \{X_{Obs}: X > 402\}
$$

```{r}
qbinom(0.95, 759, 0.50)
```

5.  What decision can be made? </br> </br> If the null hypothesis *is
    true*, there is a 0.0004 probability that another random sample of
    $n = 759$ Millennial Canadians will yield a number who are relying
    on an inheritance to achieve their financial goals that **is greated
    than the observed number/value** of $X_{Obs} = 425$. Such an event -
    $X_{Obs} > 425$ - is not likely *if the null hypothesis is true*. As
    a result, we have distrust in that assumption and the **null
    hypothesis is rejected**. </br> </br> As a result we conclude **from
    these data/this sample**\* that $p > 0.50$, or that more than 50% of
    Millennial Canadians rely on a financial inheritance to achieve
    their financial goals.

------------------------------------------------------------------------

# Testing About the Population Mean $\mu$

**Time to Play With Data 2:** Consider your **counties.sampled** data
from last Tuesday's class, where you created a 95% confidence interval
for $\mu_{Median.Earnings.2010}$. If possible, refer to your
counties.sampled data frame.

From your sample of $n = 50$ counties, can you conclude that the
mean/average median income of all $N = 3143$ counties in the United
States is more than 26,500?

**Answer:** The statistical hypothesis is $$
{\rm H}_{0}: \mu_{Median.Earnings.2010} \leq 26000 \hspace{0.5in} {\rm H}_{A}: \mu_{Median.Earnings.2010} > 26000
$$

Here is my **counties.sampled** from last Tuesday:

```{r}
head(counties.sampled, 4)
tail(counties.sampled, 4)
```

We have already inspected a Normal Probability plot of these data....in
my case, it appeared to be Normal (for the most part)....

Now that these data are collected, the sample statistics are computed
with `favstats`:

```{r}
favstats(~Median.Earnings.2010, data = counties.sampled)
```

The test statistic is computed with $$
T_{Obs} = \frac{\overbrace{26156.15}^{\overline{X}} - \overbrace{26000}^{\mu_{0} = 26000}}{\frac{S = 6883.299}{\sqrt{50}}} = 
$$

```{r}
mu.0 <- 26000
dof <- favstats(~Median.Earnings.2010, data = counties.sampled)$n - 1 #set the degrees of freedom to n - 1
t.obs <- (favstats(~Median.Earnings.2010, data = counties.sampled)$mean - mu.0)/(favstats(~Median.Earnings.2010, data = counties.sampled)$sd /sqrt(favstats(~Median.Earnings.2010, data = counties.sampled)$n))
t.obs
```

which produces $T_{Obs} = 0.1604$,

The $P$-value is $P(T_{49} > T_{Obs}) = P(T_{49} > 0.1604)$, which is
computed with R below.

```{r}
1 - pt(0.1604, dof)
```

and $$
P-\text{value} = P(T_{49} > 0.1604) = 0.4366127
$$

This can also be obtained with `t.test`:

```{r}
t.test(~Median.Earnings.2010, mu = 26000, alternative = "greater", data=counties.sampled) # use $stat to obtain the value of T_{Obs} only
```

------------------------------------------------------------------------

**Time to Play With Data 3:**

```{r}
oct4ttp3.df <- read.csv("https://people.ucalgary.ca/~jbstall/DataFiles/SalesComparison.csv")
```

```{r}
oct4ttp3.df
```

Given that these data were collected using a **Matched Pairs
Experimental Design**, the difference is defined as

$$
d_{i} = x_{Campaign, i} - x_{Without\_Campaign, i}
$$

the statistical hypothesis to be tested is

$$
\begin{eqnarray}
{\rm H}_{0}: \mu_{d} & \leq  (= ) & 0 \hspace{0.2in} \text{(the mean of a population of differences is at most equal to zero)} \\
{\rm H}_{A}: \mu_{d} & > &  0 \hspace{0.2in} \text{(the mean of a population of differences is at most equal to zero)} \\
\end{eqnarray}
$$

Adding a new variable to the **oct4ttp3.df**

```{r}
oct4ttp3.df <- oct4ttp3.df %>%
  mutate(DiffWeeklySales = (With.Campaign - Without.Campaign))
```

```{r}
oct4ttp3.df
```

```{r}
ggplot(oct4ttp3.df, aes(sample = DiffWeeklySales)) + stat_qq(col="blue") + stat_qqline(col="red") + ggtitle("Normal Probability Plot of Differences")
```

We are computing $$
T_{Obs} = \frac{\overline{d} - \mu_{d}}{\frac{s_{d}}{\sqrt{n}}}
$$

```{r}
favstats(~DiffWeeklySales, data=oct4ttp3.df)
```

and computing $$
T_{Obs} = \frac{3.653846 - 0}{\frac{3.185544}{\sqrt{13}}} = 4.1356 \:\: \text{(computed with the t.test command below)}
$$

```{r}
t.test(oct4ttp3.df$With.Campaign, oct4ttp3.df$Without.Campaign, mu=0, paired = TRUE, alternative="greater")
t.test(~DiffWeeklySales, mu = 0, alternative="greater", data = oct4ttp3.df)
```

The $P$-value is computed as

$$
P(\overline{d} > 3.653846) = P(T_{13 - 1} > 4.1356) = 0.0006908 \approx 0.00069
$$

```{r}
1 - pt(4.1356, 13 - 1)
```

**Interpretation:** If data were collected using the same methodology as
the existing/current data, the chance that we would observe a result
that is "more surprising" - or stronger statistical evidence against the
null hypothesis than the current sample/data - is $0.00069$ \*if the
null hypothesis is true.

Because it is highly unlikely to observe stronger statistical evidence
against ${\rm H_{0}}$ than the current data, the null hypothesis is
rejected in favour of the alternative hypothesis. One can infer/conclude
*from these data* that $\mu_{d} > 0$.

A 95% confidence interval for $\mu_{d}$ is then

```{r}
t.test(~DiffWeeklySales, data=oct4ttp3.df)$conf
```

$$
1.7288 \leq \mu_{d} \leq 5.5789
$$
