---
title: "DATA 602 - Tuesday, September 13"
output: html_notebook
---

**Example 1:** Here is the data

```{r}
head(SAT_2010, 4)
tail(SAT_2010, 4)
```

```{r}
names(SAT_2010)
```

A plot of the **expenditure** to average SAT-**math** score shows the following:

```{r}
ggplot(data=SAT_2010, aes(x = expenditure, y = math)) + geom_point(col="blue", size=2) + xlab("State Expenditure per Student") + ylab("SAT - Math Score") + ggtitle("Scatterplot of Expenditure/Student to State SAT-Score (Average)")
```

What story does this visualization seem to tell?

Let's look at the proportion/percentage of high school students in the state that write the SAT.

```{r}
SAT_2010$sat_pct
```

Let's create a **categorical variable** that assigned a *class* to a percentage. That is

-   if the percentage is $0 < 33$ -\> "low"
-   if the percentage is $33 < 67$ -\> "medium"
-   if the percentage is $67 \leq 100$ -\> "high"

To do so, we will use the 'mutate' function to (i) create a new variable and (ii) 'tack' this variable onto the SAT_2010 data frame.

```{r}
SAT_2010 <- SAT_2010 %>%
  mutate(SAT_class = cut(sat_pct, breaks=c(0, 33, 67, 100), labels = c("low", "medium", "high")))
```

```{r}
head(SAT_2010, 4)
```

Computing the population mean $\mu_{AveSalary}$ is

```{r}
mean(~salary, data=SAT_2010)
```

and the distribution of the **SAT_class**:

```{r}
table(SAT_2010$SAT_class)
```

------------------------------------------------------------------------

Now, let's take a look at the same scatterplot, including the "SAT_class":

```{r}
ggplot(data=SAT_2010, aes(x = expenditure, y = math, color=SAT_class)) + geom_point(size=2) + xlab("State Expenditure per Student") + ylab("SAT - Math Score") + ggtitle("Scatterplot of Expenditure/Student to State SAT-Score (Average)")
```

We can separate these out:

```{r}
ggplot(data=SAT_2010, aes(x = expenditure, y = math, color=SAT_class)) + geom_point(size=2) + xlab("State Expenditure per Student") + ylab("SAT - Math Score") + ggtitle("Scatterplot of Expenditure/Student to State SAT-Score (Average)") + facet_wrap(~SAT_class) + geom_smooth(method = "lm")
```

**Time to Play 1:** In this exercise, we will use the 'flights' data, which consists of all departing flights from the 3-NYC airports in 2013.

```{r}
head(flights, 4)
tail(flights, 4)
names(flights)
```

(a) Compute $\mu_{distance}$ (in miles)

```{r}
mean(~distance, data=flights, rm.na=TRUE)
```

$$
\mu_{distance} = 1039.913 \ \text{miles}
$$

(b) Compute $\widetilde{\mu}_{distance}$.

```{r}
median(~distance, data=flights)
median(flights$distance)
```

```{r}
favstats(~distance, data=flights)
```

$$
\widetilde{\mu}_{distance} = 872 \: \text{miles}
$$

(c) Compute $p_{Delta}$, proportion of all flights leaving the three NYC-area airports that are Delta flights.

```{r}
table(flights$carrier)
```

```{r}
total.flights <- sum(table(flights$carrier))
total.flights
```

```{r}
prop_delta = 48110/total.flights
prop_delta
```

$$
p_{Delta} = \frac{48110}{336776} = 0.1428546 \hspace{0.2in} \text{or} \hspace{0.2in} 14.29\%
$$

**Time to Play 2:** Consider the two variables **pupil_teacher_ratio** and **salary** (average teacher salary in the state). Use **salary** as the $y$-axis....

(a) What is the relationship between these two variables?

```{r}
ggplot(SAT_2010, aes(x= pupil_teacher_ratio, y = salary)) + geom_point(col="blue", size=2) + ylab("Average Teacher Salary per State") + xlab("Students/Teacher") + ggtitle("Scatterplot of Student/Teacher Ratio and Average Salary per State") + geom_smooth(method="lm")
```

(b) Create a variable called **sal_category** which will categorize a state according to: (40K \< 50K, low), (50K \< 60K, med), (60K \< 70K, medhigh), (70K \< 80K, high)

SAT2010 \<- SAT_2010 %\>% mutate(sal_category = cut(salary, c(40, 50, 60, 70, 80), labels = c("low", "med", "medhigh", "high")))

```{r}
head(SAT_2010, 4)
```

```{r}
table(SAT_2010$sal_category)
```

The scatterplot, which takes into account the distribution of the salary category:

OR

```{r}
```

```{r}
ggplot(SAT_2010, aes(x = math, y = write)) + geom_point(size = 2, col="blue") + ylab("State: Write SAT Average") + xlab("State: Math SAT Average") + ggtitle("Scatterplot of Student/Teacher Ratio and Average Salary per State") + geom_smooth(method="lm")
```

------------------------------------------------------------------------

**Time to Play 2:**

Create a data frame called **demoairline1**

```{r}
demoairline1 <- select(flights, year, month, day, dep_delay, carrier, arr_delay, distance)
```

```{r}
head(demoairline1, 4)
tail(demoairline1, 4)
```

Visualizing the variable **dep_delay**, the number of minutes a deparing flight is delayed.

```{r}
ggplot(demoairline1, aes(x = dep_delay)) + geom_histogram(col="red", fill="blue", binwidth=20, na.rm=TRUE) + xlab("Number of Minutes a Flight is Delayed in Departing") + ylab("Count") + ggtitle("Histogram of Departure Delays from the 3-NYC Area Airports")
```

```{r}
count(is.na(demoairline1$dep_delay))
```

Alternatively, one can create a **density plot**

```{r}
ggplot(demoairline1, aes(x = dep_delay)) + geom_density(col="red", fill="blue", na.rm=TRUE) + xlab("Number of Minutes a Flight is Delayed in Departing") + ylab("Density") + ggtitle("Density Plot of Departure Delays from the 3-NYC Area Airports")
```

where the $y$-axis, or density, is

$$
\text{density}_i = \frac{\text{class count/frequency}_{i}}{\text{class width}_{i}}
$$

Or, by **carrier**

```{r}
ggplot(demoairline1, aes(x = dep_delay)) + geom_histogram(col="red", fill="blue", binwidth=20, na.rm=TRUE) + xlab("Number of Minutes a Flight is Delayed in Departing") + ylab("Count") + ggtitle("Histogram of Departure Delays from the 3-NYC Area Airports") + facet_wrap(~carrier)
```

```{r}
ggplot(demoairline1, aes(x = dep_delay)) + geom_density(col="red", fill="blue", na.rm=TRUE) + xlab("Number of Minutes a Flight is Delayed in Departing") + ylab("Density") + ggtitle("Density Plot of Departure Delays from the 3-NYC Area Airports") + facet_wrap(~carrier)
```

Suppose you wanted to filter one one airline, say **DL** and create a histogram of Delta's departure delays.

```{r}
ggplot(filter(demoairline1, carrier == "DL"), aes(x = dep_delay)) + geom_histogram(col="red", fill="blue", binwidth = 10, na.rm=TRUE) + xlab("Number of Minutes a Flight is Delayed in Departing") + ylab("Density") + ggtitle("Histogram of Departure Delays of Delta Airlines from the 3-NYC Area Airports")
```

Now, try this for United Airlines, "UA":

```{r}
ggplot(filter(demoairline1, carrier == "UA"), aes(x = dep_delay)) + geom_histogram(col="red", fill="blue", binwidth = 10, na.rm=TRUE) + xlab("Number of Minutes a Flight is Delayed in Departing") + ylab("Density") + ggtitle("Histogram of Departure Delays of United Airlines from the 3-NYC Area Airports")
```

Create a subset that consists *only of United flights*

```{r}
ua.flights <- subset(demoairline1, carrier == "UA")
```

```{r}
head(ua.flights, 4)
tail(ua.flights, 3)
```

```{r}
dim(ua.flights)
```

```{r}
ggplot(ua.flights, aes(x = dep_delay)) + geom_density(fill="blue", na.rm=TRUE) + xlab("Number of Minutes a Flight is Delayed in Departing") + ylab("Density") + ggtitle("Density Plot of Departure Delays of United Airlines from the 3-NYC Area Airports") + geom_vline(xintercept=0, col="red")
```

Now, suppose we wanted to compute the proportion of all United Flights that were "late" in departing.

```{r}
nrow(subset(ua.flights, dep_delay > 0))
```

in this instance, there are **27261** flights that are late in departing. The number of flights that have data are

```{r}
count(is.na(ua.flights$dep_delay))
```

```{r}
num_ua_flights <- 58665 - 686
```

```{r}
num_ua_flights
```

```{r}
pua_late <- nrow(subset(ua.flights, dep_delay > 0))/num_ua_flights
pua_late
```

and

$$
p_{UA, late\:in\:departing} = 0.47021875
$$

**Time to Play 1:**

use the command

```{r}
cl.limits = seq(-100, 1300, 25)
```

(a) A histogram of the arrival delays **arr_delay** for all airlines in the **demoairline1** data frame.

```{r}
ggplot(demoairline1, aes(x = arr_delay)) + geom_histogram(fill="blue", breaks = cl.limits, na.rm=TRUE) + xlab("Minutes a Flight Departing 3-NYC Area Airports is late @ Destination") + geom_vline(xintercept=0, col="red")
```

```{r}
dim(demoairline1)
count(is.na(demoairline1$arr_delay))
```

There are 9430 missing data points. Therefore

```{r}
deno <- 336776 - 9430
nrow(subset(demoairline1, arr_delay < 0))
```

```{r}
p_early <- 188933/deno
p_early
```

and

$$
p_{arrive\:\:early} = \frac{188933}{327346} = 0.5771661 \approx 0.5772
$$

(c) 

```{r}
aa.flights <- filter(demoairline1, carrier == "AA")
```

```{r}
head(aa.flights, 4)
tail(aa.flights, 4)
```

```{r}
dim(aa.flights)
count(is.na(aa.flights$arr_delay))
```

```{r}
ggplot(aa.flights, aes(x = arr_delay)) + geom_density(col="blue", na.rm=TRUE) + geom_vline(xintercept = 0, col="orange")
```

There are 31947 flights AA flights with valid data (arr_delay)

```{r}
nrow(subset(aa.flights, arr_delay > 0))
```

```{r}
10706/31947
```

$$
p_{AA, late\:in\:arriving} = \frac{10706}{31947} = 0.3351175 \approx 0.3351
$$

**Time to Play 2:**

(a) Relationship between **distance** and **arr_delay**?

```{r}
ggplot(demoairline1, aes(x = distance, y = arr_delay)) + geom_point(col="blue", position="jitter", size=1, na.rm=TRUE) + ylab("Minutes the Flight is Late in Arriving @ Destination") + xlab("Distance in Miles") + ggtitle("Scatterplot of Arrival Time to Distance") + geom_smooth(method="lm")                                                                                                    
```

(b) Create a similar visualization for (i) Delta (DL) and (ii) United Airlines (UA)

```{r}
ggplot(filter(demoairline1, carrier == "UA"|carrier == "DL"), aes(x = distance, y = arr_delay, color=carrier)) + geom_point(position = "jitter", na.rm=TRUE) + xlab("Flight Distance in Miles") + ylab("Minutes Late in Arriving at Destination") + ggtitle("Scatterplot: Distance of Flight and Minutes Late in Arrriving")
```
