---
title: "R Notebook"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
  word_document: default
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(binom)
library(car)
library(collapsibleTree)
library(dbplyr)
library(dplyr)
library(EnvStats)
library(ggformula)
library(ggplot2)
library(gmodels)
library(htmltools)
library(ISLR)
library(knitr)
library(lawstat)
library(markdown)
library(mosaic)
#library(mdsr)
library(mosaicData)
library(nycflights13)
library(olsrr)
library(plyr)
library(purrr)
library(plotly)
library(resampledata)
library(rmarkdown)
library(rpart)
library(rpart.plot)
library(rvest)
library(SDaA)
library(shiny)
library(stringi)
library(tibble)
library(tidyr)
library(tidyselect)
library(tinytex)
library(yaml)
library(shiny)

```

# DATA 602 PROJECT

Zheyu (Jerry) Song, Hao (Leo) Su, Kelly Wu (Group 11)

# Introduction & Background:

For our project we explored the relationship between COVID-19 Pandemic
outcomes, such as positive cases of COVID-19 and related deaths to
COVID-19 related to obesity.

We are interested in obesity as a factor in COVID-19 related outcomes
because obesity is a non-communicable chronic disease that can impact
one's overall health and their ability to defend themselves against
other illnesses, such as COVID-19. Already, a study from 2021 on the USA
population found that the the risk for severe COVID-19 related outcomes
increased with higher BMI (Kompaniyets et al., 2021).

Before we look into the relationship between COVID-19 outcomes and
obesity directly, we wish to understand the relationship between Gross
Domestic Product (GDP) and COVID-19. The reason for looking into the
GDP's relationship with obesity, is to understand if the type of
country, i.e. developed/developing matters when looking at COVID-19
outcomes and the the obesity rate. In early 2022, a research paper by
Oshakbayev et al., discussed findings that of analyses that showed
significant correlations between GDP and obesity, and overweight
prevalence (2022).

## Datasets Used:

Overweight and obesity based on measured body mass index, by age group
and sex (Open Government, 2021)

COVID-19 Healthy Diet Dataset (Kaggle, 2020)

Cleaned data set from DATA 601 (dervied from the COVID-19 Healthy Diet
Dataset)

## Guiding Questions

Question 1:

Firstly, we analyze the trends of obesity in Canada and it's
relationship with Canada's Gross Domestic Product (GDP) value. We will
do this using statistical hypotheses.

Question 2:

Secondly, we will analyze and model the relationship between obesity and
each COVID-19 outcome, deaths as a result of COVID-19 and confirmed
positive cases of COVID-19. We will do this using linear regression and
bootstrapping methods.

# Reading the data

```{r}
food = read.csv("Food_Supply_Quantity_kg_Data.csv")
canada = read.csv("13100373.csv")
processed = read.csv("haha.csv")
head(food, 3)
head(canada, 3)
head(processed, 3)
```

# Data wrangling

```{r}
food <- select(food, c('Country', 'Obesity', 'Undernourished', 'Confirmed', 'Deaths', 'Recovered'))
food <- na.omit(food)
head(food, 3)
```

```{r}
canada <- select(canada, c('REF_DATE', 'Measures', 'Sex', 'Age.group', 'Characteristics', 'VALUE'))
head(canada, 3)
```

```{r}
canada.percent <- filter(canada, Measures == 'Obese' & Sex == 'Both sexes' & Age.group == 'Ages 5 to 79' & Characteristics == 'Percent')

canada.number <- filter(canada, Measures == 'Obese' & Sex == 'Both sexes' & Age.group == 'Ages 5 to 79' & Characteristics == 'Number of persons')

head(canada.percent, 3)
head(canada.number, 3)
```

```{r}
processed <- select(processed, c('Country', 'Obesity', 'Undernourished', 'Confirmed', 'Deaths', 'region'))
head(processed,3)
```

# Visualizing the overall data from our data sets

## Violin Plots to Demonstrate the Distribution of countries for each variable: Obesity Rate, Confirmed COVID-19 Positive Cases, Deaths as a result of COVID-19

```{r}
ggplot(data = food, aes(x = "All Sample", y = Obesity)) + geom_violin(col="blue", fill="red", na.rm=TRUE) + geom_boxplot(width=0.1, na.rm=TRUE) + xlab("Countries") + ylab("The Obesity Rate") + ggtitle("Violin Plot of the Obesity Rate") + coord_flip()

ggplot(data = food, aes(x = "All Sample", y = Confirmed)) + geom_violin(col="blue", fill="red", na.rm=TRUE) + geom_boxplot(width=0.1, na.rm=TRUE) + xlab("Countries") + ylab("The Confirmed COVID-19 Cases Rate") + ggtitle("Violin Plot of the Confirmed COVID-19 Cases Rate") + coord_flip()

ggplot(data = food, aes(x = "All Sample", y = Deaths)) + geom_violin(col="blue", fill="red", na.rm=TRUE) + geom_boxplot(width=0.1, na.rm=TRUE) + xlab("Countries") + ylab("The Deaths Related to COVID-19 Rate") + ggtitle("Violin Plot of the Deaths Related to COVID-19 Rate") + coord_flip()
```

### Describing the Violin Plots

At initial glance, we can see that many countries have low rates of
COVID-19 related deaths and confirmed cases. We may attribute this to
the large number of smaller countries in our data set that would have
smaller populations and in turn, have lower COVID-19 related outcomes
(deaths and positive cases).

Wider sections of the violin plot represent a higher probability that
members of the population will take on the given value Looking at the
violin plot that shows the obesity rate, we see that there is a higher
probability that countries will have an obesity rate that is greater
than the median of the sample.

```{r}
obesitymean=aggregate(processed$Obesity, by=list(type=processed$region),mean)
obesitymean

undernourishedmean=aggregate(processed$Confirmed, by=list(type=processed$region),mean)
undernourishedmean

confirmedmean=aggregate(processed$Confirmed, by=list(type=processed$region),mean)
confirmedmean

deathsmean=aggregate(processed$Deaths, by=list(type=processed$region),mean)
deathsmean
```

## Bar Graphs to Demonstrate the Distribution of Each Variable: Obesity Rate, Confirmed COVID-19 Positive Cases, Deaths as a result of COVID-19 Over the Number of Countries

```{r}
ggplot(data=obesitymean, aes(x=reorder(type,x), y=x, fill=type)) + geom_bar(stat="identity") + xlab("Region") + ylab("The Obesity Rate") + ggtitle("Bar Graph of the Obesity Rate in Each Continent")

ggplot(data=confirmedmean, aes(x=reorder(type,x), y=x, fill=type)) + geom_bar(stat="identity") + xlab("Region") + ylab("The Confirmed COVID-19 Cases Rate") + ggtitle("Bar Graph of the Confirmed COVID-19 Cases Rate in Each Continent")

ggplot(data=deathsmean, aes(x=reorder(type,x), y=x, fill=type)) + geom_bar(stat="identity") + xlab("Region") + ylab("The Deaths Related to COVID-19 Rate") + ggtitle("Bar Graph of the Deaths Related to COVID-19 Rate in Each Continent")
```

#### Describing the Bar Graphs

As you can see, the bar graphs showing the amount of COVID-19 related
deaths and confirmed positive cases of COVID-19 are very similar,
including the order of the countries. A large gap can also be found
between the Europe and the next continent.

In the graph showing the obesity rate among continents, Oceania leads
with the highest rate of obesity, followed by Europe, Americas, Asia and
then Africa. The gaps between these continents are much smaller for this
graph than the others.

## Simple Line Graph to Demonstrate the obesity rate in Canada between the years of:

```{r}
ggplot(canada.percent, aes(x=REF_DATE, y=VALUE)) + geom_point(color="red", size=3) + geom_line(color="blue", size=1) + xlab("Years") + ylab("The Obesity Rate of Canada") + ggtitle("Line Graph of the Obesity Rate of Canada")
```

### Describing the Line Graph

The line graph visualizes the trend of obesity in Canada between the
years: 2009 and 2019.

# Question 1: Analyze the trends of obesity in Canada and it's relationship with Canada's Gross Domestic Product (GDP) value.

### Part A: Compare obesity rate for years 2009 and 2013

Reasoning for the years chosen: Canada's GDP per capita is the lowest
after the 2008 recession and Canada's GDP in 2013 GDP is the second
highest after 2008 recession. We also chose the year 2013 becuase our
data set has data for that year.

A a random selection of $n_{2009} = 6479500$ Canadians aged 5 to 79
years old, of which $22.4%$ of the sample size are obese. A similar poll
by the Government of Canada in 2013 had a sample size of
$n_{2013} = 7424000$ of Canadians aged 5 to 79 years old found that
$24.2%$ of the sample size are obese.

We want to know if this data indicates that the obesity rate in Canada
in 2009 is higher than the obesity rate in 2013. Carry out the
statistical test.

The statistical hypotheses to be tested are:

$$
\begin{eqnarray}
{\rm H}_{0}: p_{Obe\_2013} & \leq  & p_{Obe\_2009} \hspace{0.2in} \text{(The obesity rate in 2013 is not higher than the obesity rate in 2009)} \\
{\rm H}_{A}: p_{Obe\_2013} & >  & p_{Obe\_2009} \hspace{0.2in} \text{(The obesity rate in 2013 is higher than the obesity rate in 2009)} \\
\end{eqnarray}
$$

We set up the $\alpha = 0.05$

Next we compute the value of the test statistic.

```{r}
p_hat2009 = 0.224
p_hat2013 = 0.242
n2009 = 6479500 #sample size selected in 2009
x2009 = n2009 * p_hat2009 #the number of random selected Canadians aged 5 to 79 years old who is obese in 2009
n2013 = 7424000 #sample size selected in 2013
x2013 = n2013 * p_hat2013 #the number of random selected Canadians aged 5 to 79 years old who is obese in 2013
x2009
x2013
```

In this computation, we assume a "common proportion" under
${\rm H}_{0}$. The estimate of this common proportion, the "pooled
sample proportion" $\widehat{p}$ is:

```{r}
pooled.p1 <- (x2009 + x2013)/(n2009 + n2013)
pooled.p1
```

$$
\widehat{p} = \frac{X_{Obe\_2009} + X_{Obe\_2013}}{n_{Obe\_2009} + n_{Obe\_2013}} = \frac{1451408 + 1796608}{6479500 + 7424000} = 0.2336114
$$ Compute the test statistic $Z_{Obs}$

$$
Z_{obs} = \frac{\widehat{p}_{Obe\_2013} - \widehat{p}_{Obe\_2009} - (p_{Obe\_2013} - p_{Obe\_2009})}{\sqrt{\widehat{p}(1 - \widehat{p})\left(\frac{1}{n_{Obe\_2013}} + \frac{1}{n_{Obe\_2009}} \right)}} = \frac{0.242 - 0.224 - (0)}{\sqrt{0.2336114(1-0.2336114)\left(\frac{1}{7424000} + \frac{1}{6479500} \right)}} = 79.12772
$$

```{r}
z.obsnumerator1 <-  p_hat2013 - p_hat2009
z.obsdenominator1 <- sqrt(pooled.p1 * (1 - pooled.p1)*((1/n2009) + (1/n2013)))
z.obs1 <- z.obsnumerator1/z.obsdenominator1

z.obs1
```

```{r}
1-pnorm(z.obs1)
```

The $P$-value is then $P(Z >79.12772)$ which is computed to be $0$

Because our $p-value < \alpha$, Reject $H_{0}$. We conclude that The
obesity rate in 2013 is higher than the obesity rate in 2009.

The computation of the test statistic above can be completed with the
`prop.test` command.

```{r}
prop.test(c(x2013,x2009), c(n2013,n2009), alternative="greater", correct=FALSE)
```

### Part B: Compare obesity rate for years 2013 and 2019

Comparing the biggest gap between the lowest point with the highest
point before the COVID-19 pandemic.

We want to know if these data indicate that the obesity rate in 2013 is
higher than the obesity rate in 2019.

A a random selection of $n_{2019} = 7169800$ Canadians aged 5 to 79
years old, of which $22.1%$ the sample size are obese. A similar poll by
the Government of Canada in 2013 had a sample size of
$n_{2013} = 7424000$ Canadians aged 5 to 79 years old found that $24.2%$
of the sample size are obese in 2013.

The statistical hypothesese to be tested are:

$$
\begin{eqnarray}
{\rm H}_{0}: p_{Obe\_2013} & \leq  & p_{Obe\_2009} \hspace{0.2in} \text{(The obesity rate in 2019 is not higher than the obesity rate in 2013)} \\
{\rm H}_{A}: p_{Obe\_2013} & >  & p_{Obe\_2009} \hspace{0.2in} \text{(The obesity rate in 2019 is higher than the obesity rate in 2013)} \\
\end{eqnarray}
$$

We set up the $\alpha = 0.05$ Next we compute the value of the test
statistic.

```{r}
p_hat2019 = 0.221
p_hat2013 = 0.242
n2019 = 7169800 #sample size selected in 2009
x2019 = n2019 * p_hat2019 #the number of random selected Canadians aged 5 to 79 years old who is obese in 2009
n2013 = 7424000 #sample size selected in 2013
x2013 = n2013 * p_hat2013 #the number of random selected Canadians aged 5 to 79 years old who is obese in 2013
x2019
x2013
```

In this computation, we assume a "common proportion" under
${\rm H}_{0}$. The estimate of this common proportion, the "pooled
sample proportion" $\widehat{p}$ is:

```{r}
pooled.p2 <- (x2019 + x2013)/(n2019 + n2013)
pooled.p2
```

$$
\widehat{p} = \frac{X_{Obe\_2019} + X_{Obe\_2013}}{n_{Obe\_2019} + n_{Obe\_2013}} = \frac{1451408 + 1796608}{6479500 + 7424000} = 0.2336114 
$$ Compute the test statistic $Z_{Obs}$

$$
Z_{obs} = \frac{\widehat{p}_{Obe\_2013} - \widehat{p}_{Obe\_2019} - (p_{Obe\_2013} - p_{Obe\_2019})}{\sqrt{\widehat{p}(1 - \widehat{p})\left(\frac{1}{n_{Obe\_2013}} + \frac{1}{n_{Obe\_2019}} \right)}} = \frac{0.242 - 0.224 - (0)}{\sqrt{0.2336114(1-0.2336114)\left(\frac{1}{7424000} + \frac{1}{6479500} \right)}} = 79.12772
$$

```{r}
z.obsnumerator2 <-  p_hat2013 - p_hat2019
z.obsdenominator2 <- sqrt(pooled.p2 * (1 - pooled.p2)*((1/n2019) + (1/n2013)))
z.obs2 <- z.obsnumerator2/z.obsdenominator2

z.obs2
```

```{r}
1-pnorm(z.obs2)
```

The $P$-value is $P(Z >79.12772)$ which is computed to be $0$.

Because $p-value < \alpha$, Reject $H_{0}$. We conclude that The obesity
rate in 2013 is higher than the obesity rate in 2019.

The computation of the test statistic above can be completed with the
`prop.test` command.

```{r}
prop.test(c(x2013,x2019), c(n2013,n2019), alternative="greater", correct=FALSE)
```

### Part C: Is there ample statistical evidence to confirm that, in the year 2019, the proportion of Canadians aged 5 to 79 years old who were obese, was less than 31.3%?

From the "Overweight and obesity based on measured body mass index, by
age group and sex" (Open Government, 2021), we know that out of the
random sample of $n_{2019} = 7169800$ Canadians aged 5 to 79 years old,
1584526 were obese in 2019. Compared to the "COVID-19 Healthy Diet
Dataset" (Kaggle, 2021) obesity values taken from the year 2021, 31.3%
of Canadians were obese.

$$
\begin{eqnarray}
{\rm H}_{0}: p  & \ge & \:0.313 \hspace{0.25in} {\rm \text{(the proportion of Canadians aged 5 to 79 years old who are Obese, is not less than 31.3%,in 2019)}} \\
{\rm H}_{A}: p & <& \:0.313 \hspace{0.25in} {\rm \text{(the proportion of Canadians aged 5 to 79 years old who are Obese, is less than 31.3%, in 2019)}} \\
\end{eqnarray}
$$

Set up $\alpha = 0.05$.

The data we are looking at has the sample size of $n_{2019} = 7169800$
From this same data, the observed number who are obese is
$X_{Obs} = 1584526$, which is the test statistic. The distribution of
$X$, which can be modeled by the Binomial probability model is provided
below and is conditional upon the null hypothesis being true or
$p = 0.313$.

If the null hypothesis is true, then the distribution of $X$ will look
like this:

The Distribution of the Statistic $X_{Obs}$:

```{r}


plot(0:716, dbinom(0:716, 716, 0.313), xlab="Number of Obese Observed in n=7169800", ylab="P(X = X_{Obs})", type="h", col='red', main="Distribution of Test Statistic")
abline(v = 158, col="blue")
abline(v = 716 * 0.313 , col="black")
```

And the resulting mean/expected value is: $E(X) = \mu_{X} = 2244147$
with a standard deviation of $SD(X) = \sigma_{X} = 1241.664$.

```{r}
ex = 7169800 * 0.313
sd = sqrt(7169800 *0.313*(1-0.313) )

ex
sd
```

**Compute the** $P$-value.

$$
\begin{aligned}
P-{\rm value} = & P(X < \overbrace{1584526}^{X_{Obs}}|p = 0.313) \\
              = & \sum_{x = 0}^{1584525}{7169800 \choose x}\left(0.313\right)^{x}\left(0.687\right)^{7169800 - x} \\
              = & 0 \:\: \text{(see R code below)}\\
\end{aligned}
$$

```{r}
pbinom(1584525, 7169800, 0.313)
```

Our $P$-value is $0$, which means there is no trust in the null
hypothesis (and $P$-value of $0 < 0.05$). As a result, we decide to
reject the null hypothesis and conclude from this data that the
proportion of Canadians aged 5 to 79 years old who are obese is less
than 31.3%, in 2019.

# Question 2:

## Part A: Analyze and model the relationship between obesity and each COVID-19 outcome, deaths as a result of COVID-19 and confirmed positive cases of COVID-19.

## Visualization of The Obesity Rate and The Confirmed Rate

Create scatterplot of **The Obesity Rate** to **The Confirmed** Rate:

```{r}
ggplot(food, aes(x = Obesity, y = Confirmed)) + geom_point(col="blue", size = 2) + xlab("% of The Obesity Rate") + ylab("% of The Confirmed COVID-19 Cases Rate") + ggtitle("% The Obesity Rate to % The Confirmed COVID-19 Cases Rate") + geom_smooth(method="lm", col="red")
```

## Quantifying the Relationship between The Obesity Rate and The Confirmed Rate by Using the Correlation Coefficient

```{r}
cor(~Obesity, ~Confirmed, data=food)
```

The correlation coefficient between the obesity Rate and the confirmed
positive COVID-19 cases rate is 0.5248714. This shows that there is a
strong positive linear relationship between the rate of obesity and
confirmed positive COVID-19 cases.

```{r}
favstats(~ Obesity, data=food)
```

```{r}
favstats(~ Confirmed, data=food)
```

From this output, $$
\overline{X} = 18.59808 \hspace{0.2in} S_{X} = 9.549116 \hspace{0.2in} \overline{Y} = 2.0719 \hspace{0.2in} S_{Y} = 2.385296  \hspace{0.2in} r = 0.5248714
$$ \## Modeling the Relationship Between The Obesity Rate and The
Confirmed Rate

The *statistical model* is:

$$
\underbrace{Confirmed Rate_{i}}_{Y-variable} = A + (B*\underbrace{Obesity Rate_{i}}_{X-variable}) + e_{i}
$$

We use the values of $r$, $S_{X}$, and $S_{Y}$ computed.

$$
b = r\left( \frac{S_{Y}}{S_{X}} \right) = (0.5248714)\left(\frac{2.385296}{9.549116}\right) \approx 0.1311088 
$$

We now compute the value of $a$: $$
a = \overline{Y} - (b*\overline{X}) = 2.0719 - (0.1311088*18.59808) \approx  -0.3664725 
$$

and the estimate of the model is then

$$
\widehat{Confirmed Rate}_{i} = -0.3664725  + 0.1311088*Obesity Rate_{i}
$$

Compute the estimate with R using the `lm` function.

```{r}
predictdemovote = lm(Confirmed ~ Obesity, data=food)  
```

```{r}
options(scipen=999)
predictdemovote$coef
```

The value of $a$ and $b$ ar $a = -0.3664725$ and $b = 0.1311088$. From
the training data, we have an estimate of the model.

$$
\widehat{Confirmed Rate}_{i} = -0.3664725 + (0.1311088*Obesity Rate_{i})
$$

Notes about the values A & B:

A: when the obesity rate is 0, the confirmed rate is -0.3664725%. This
is meaningless because most countries do not have an obesity rate of 0.

B: As the obesity rate increases by 1%, then the confirmed rate will
increase by an *average* of 0.1311088%.

### The $F$-Test of Linear Appropriateness

Checking if the linear model meets the condition of linear
appropriateness and if
$Confirmed Rate_{i} = A + (B*Obesity Rate_{i}) + e_{i}$ is a valid model
by checking whether the slope term $\beta_{1}$ is non-zero or not.

Test the statistical hypotheses: $$
{\rm H}_{0}: B = 0\:\:(Confirmed Rate\:\:{\rm cannot\:\:be\:\:expressed\:\:as\:\:a\:\:linear\:\:function\:\:of\:\:}Obesity Rate) \\
{\rm H}_{A}: B \ne 0\:\:(Confirmed Rate\:\:{\rm can\:\:be\:\:expressed\:\:as\:\:a\:\:linear\:\:function\:\:of\:\:}Obesity Rate)
$$

Set up the $\alpha = 0.05$. F-Test of Linear Appropriateness

```{r}
summary(aov(predictdemovote))
```

p-value = 0.00000000000201. $p-value < \alpha$, Due to the p-value being
below the alpha value, we can reject the null hypothesis. We can
conclude that the confirmed positive cases of COVID-19 rate can be
expressed as a linear function of the Obesity Rate.

### Normality of the Residuals of the linear model

Checking if the model meets the condition of normality of the residuals.

```{r}
predicted.values.demovote = predictdemovote$fitted.values
eisdemovote = predictdemovote$residuals
diagnosticdf = data.frame(predicted.values.demovote, eisdemovote)
head(diagnosticdf)
```

```{r}
ggplot(diagnosticdf, aes(sample = eisdemovote)) +  stat_qq(col='blue') + stat_qqline(col='red') + ggtitle("Normal Probability Plot of Residuals")
```

The plot shows that the residuals are normally distributed and meets the
condition.

### Homoscedasticity of the Residuals of the linear model

Checking if the linear model meets the condition of homoscedasticity.

```{r}
ggplot(diagnosticdf, aes(x = predicted.values.demovote, y = eisdemovote)) +  geom_point(size=2, col='blue', position="jitter") + xlab("Predicted % The Confirmed Rate") + ylab("Residuals") + ggtitle("Plot of Fits to Residuals") + geom_hline(yintercept=0, color="red", linetype="dashed")
```

Based on our plot above, we can say that the homoscedasticity condition
is satisfied.

### Analysis of Error and Coefficient of Determination of The Linear Model

```{r}
aov(predictdemovote)
```

From this output,

$$
SSE = 638.9409 \hspace{0.2in} SSR = 242.9528   \hspace{0.2in} df = 154 \hspace{0.2in} SST = SSE + SSR = 881.8937  \hspace{0.2in} S_{e} = 2.036901
$$

**Coefficient of Determination**

```{r}
rsquared(predictdemovote)
```

The coefficient of determination is 0.2754899. This coefficient of
determination is too small, which means that our model is not good.

## Part B:

### Visualizing of The Obesity Rate and The Deaths

Scatterplot of **The Obesity Rate** to **The COVID-19 Related Deaths
Rate** Rate:

```{r}
ggplot(food, aes(x = Obesity, y = Deaths)) + geom_point(col="blue", size = 2) + xlab("% of The Obesity Rate") + ylab("% of The Deaths Related to COVID-19 Rate") + ggtitle("% The Obesity Rate to % The Deaths Related to COVID-19 Rate") + geom_smooth(method="lm", col="red")
```

### The $F$-Test of Linear Appropriateness

Checking if the linear model meets the condition of linear
appropriateness and if
$Deaths Rate_{i} = A + (B*Obesity Rate_{i}) + e_{i}$ is a valid model by
checking whether the slope term $\beta_{1}$ is non-zero or not.

Test the statistical hypotheses:

$$
{\rm H}_{0}: B = 0\:\:(Deaths Rate\:\:{\rm cannot\:\:be\:\:expressed\:\:as\:\:a\:\:linear\:\:function\:\:of\:\:}Obesity Rate) \\
{\rm H}_{A}: B \ne 0\:\:(Deaths Rate\:\:{\rm can\:\:be\:\:expressed\:\:as\:\:a\:\:linear\:\:function\:\:of\:\:}Obesity Rate)
$$ Set up the $\alpha = 0.05$.

```{r}
predictdemovote = lm(Deaths ~ Obesity, data=food)
summary(aov(predictdemovote))
```

p-value = 0.00000000000201 $p-value < \alpha$. Since our p-value is less
than our alpha value, we can reject the null hypothesis. We can conclude
that the COVID-19 related deaths rate can be expressed as a linear
function of the obesity rate.

### Normality of the Residuals of the linear model\*\*

```{r}
predicted.values.demovote = predictdemovote$fitted.values
eisdemovote = predictdemovote$residuals
diagnosticdf = data.frame(predicted.values.demovote, eisdemovote)
head(diagnosticdf)
```

```{r}
ggplot(diagnosticdf, aes(sample = eisdemovote)) +  stat_qq(col='blue') + stat_qqline(col='red') + ggtitle("Normal Probability Plot of Residuals")
```

The plot shows that the residuals are normally distributed.

### Homoscedasticity of the Residuals of the linear model

```{r}
ggplot(diagnosticdf, aes(x = predicted.values.demovote, y = eisdemovote)) +  geom_point(size=2, col='blue', position="jitter") + xlab("Predicted % The Deaths Rate") + ylab("Residuals") + ggtitle("Plot of Fits to Residuals") + geom_hline(yintercept=0, color="red", linetype="dashed")
```

Based on our graph above we can conclude that the homoscedasticity
condition is satisfied.

### Modeling the Relationship Between The Obesity Rate and the COVID-19 Related Deaths Rate by Bootstrapping

```{r}
Nbootstraps = 1000 #resample n =  200, 1000 times
cor.boot = numeric(Nbootstraps) #define a vector to be filled by the cor boot stat
a.boot = numeric(Nbootstraps) #define a vector to be filled by the a boot stat
b.boot = numeric(Nbootstraps) #define a vector to be filled by the b boot stat
```

```{r}
nsize = dim(food)[1]  #set the n to be equal to the number of bivariate cases, number of rows

#start of the for loop
for(i in 1:Nbootstraps)
{   #start of the loop
    index = sample(nsize, replace=TRUE)  #randomly picks a number between 1 and n, assigns as index
    demovote.boot = food[index, ] #accesses the i-th row of the regressionex1.df data frame
    
    cor.boot[i] = cor(~Obesity, ~Deaths, data=demovote.boot) #computes correlation for each bootstrap sample
    votedemocrat.lm = lm(Deaths ~ Obesity, data=demovote.boot)  #set up the linear model
    a.boot[i] = coef(votedemocrat.lm)[1] #access the computed value of a, in position 1
    b.boot[i] = coef(votedemocrat.lm)[2] #access the computed valeu of b, in position 2

}
#end the loop
#create a data frame that holds the results of teach of he Nbootstraps 
bootstrapresultsdf = data.frame(cor.boot, a.boot, b.boot)
```

```{r}
head(bootstrapresultsdf, 3)
```

### Bootstrap Distribution of $r_{boot}$

***Coefficient of Determination***

```{r}
ggplot(bootstrapresultsdf, aes(x = cor.boot)) + geom_histogram(col="red", fill="blue", binwidth=0.025) + xlab("Values of the Bootstrap Statistic: Correlation Coefficient") + ylab("Count") + ggtitle("Distribution of Bootstrap Statistics: r")
```

```{r}
favstats(~cor.boot, data=bootstrapresultsdf)
```

```{r}
qdata(~cor.boot, c(0.025, 0.975), data=bootstrapresultsdf)
rsquared = (favstats(~cor.boot, data=bootstrapresultsdf)$mean)^2
rsquared
```

The mean of all $r_{boot}$ is 0.4933598. with a 95% confidence interval
\$0.3998882 \leq r\_{boot} \leq 0.5965969 \$ The coefficient of
determination is 0.2436653.

### Bootstrap Distribution of $a_{boot}$

```{r}
ggplot(bootstrapresultsdf, aes(x = a.boot)) + geom_histogram(col="red", fill="blue", binwidth=0.001) + xlab("Values of the Bootstrap Statistic: y-Intercept Estimate") + ylab("Count") + ggtitle("Distribution of Bootstrap Statistics: a")
```

```{r}
qdata(~a.boot, c(0.025, 0.975), data=bootstrapresultsdf)
```

```{r}
favstats(~a.boot, data=bootstrapresultsdf)
```

The mean of all $a_{boot}$ is -0.006948612 with a 95% confidence
interval \$-0.01606346 \leq a\_{boot} \leq 0.00181240 \$

### Bootstrap Distribution of $b_{boot}$

```{r}
ggplot(bootstrapresultsdf, aes(x = b.boot)) + geom_histogram(col="red", fill="blue", binwidth=0.0001) + xlab("Values of the Bootstrap Statistic: Slope Estimate") + ylab("Count") + ggtitle("Distribution of Bootstrap Statistics: b")
```

```{r}
favstats(~b.boot, data=bootstrapresultsdf)
```

```{r}
qdata(~b.boot, c(0.025, 0.975), data=bootstrapresultsdf)
```

The mean of all $b_{boot}$ is 0.002547277. with a 95% confidence
interval $0.001901773 \leq b_{boot} \leq 0.003281046$.

Using the means of $a_{boot}$ and $b_{boot}$, our estimate of the model
is:

$$
\widehat{Deaths Rate}_{i} = -0.006948612 + (0.002547277*Obesity Rate_{i})
$$

A: when the obesity rate is 0, the deaths rate is -0.006948612%. This is
meaningless because most countries do not have an obesity rate of 0.

B: As the obesity rate increases by 1%, then the deaths rate will
increase by an *average* of 0.002547277%.

# Conclusions

We can conclude that the obesity rate in Canada is higher in 2019 than
2013, and higher in 2013 than in 2009. Contrary to some economical
theories and ideas that suggest that lower GDP results in poorer eating
habits (fast food, junk food, etc.), we found higher GDP rates in Canada
were related to higher obesity prevalence. However, the study mentioned
in the introduction also found that obesity was actually more prevalent
in countries that had a higher GDP. This suggests that our findings are
in agreement with some past findings (Oshakbayev et al., 2022).

We can also conclude that there is a positive linear relationship
between obesity and COVID-19 related deaths and between obesity and
confirmed COVID-19 positive cases. Furthermore, we were also able to
model this relationship via linear regression. This supports previous
findings in a study by Kompaniyets et al. which found COVID-19 death
rates were 10 times higher in countries where more than half of the
adult population is classified as overweight (2020).

Finally, we can conclude that obesity is definitely an important factor
when looking at the outcomes of COVID-19 related outcomes.

# Limitations and Future Directions

There are some limitations to our project analysis which I will discuss
in this section and some future directions we can take based on these
limitations.

For first question, we use the variable GDP as our indicator of how well
the population is doing economically. However, this may not be the best
indicator of how well the people are doing and it may not capture how
well pecific populations are doing. In fact, in a study by Templin et
al., they looked at obesity's relationship with one's household income
and found that obesity was most prevalent among the poor (2019). We may
want to consider different economical variables to measure "how well" a
population is doing.

Additionally, because we only looked at Canada when analyzing the
relationship between GDP and obesity, we do not if this relationship
holds for other countries. It may be beneficial to look at different
countries and compare countries obesity rates by their GDP value (group
them as a developed/developing country).

Lastly for question one, our Canadian sample size included people from
ages of 5-79, which is a big age gap for our sample. This does not allow
us to look at specific age populations within the country, which
possibly could skew data.

For the second question, although we found a relationship found between
COVID-19 deaths and confirmed positive cases, we cannot specifically say
if those who were obese were the ones who had a positive case of
COVID-19 or died as a result of COVID-19. Deeper and further analysis
will be needed to determine if this relationship still exists for
specific populations.

# References

Kompaniyets, L., Goodman, A.B., Belay, B., Freedman, D.S., Sucosky,
M.S., Lange, S.J., Gundlapalli, A.V., Boehmer, T.K. and Blanck, H.M.,
2021. Body mass index and risk for COVID-19--related hospitalization,
intensive care unit admission, invasive mechanical ventilation, and
death---United States, March--December 2020. Morbidity and Mortality
Weekly Report, 70(10), p.355.

Open Government. (2021) Government of Canada. Overweight and obesity
based on measured body mass index, by age group and sex, open.canada.ca,
accessed 17 October 2022.

Oshakbayev, K., Zhankalova, Z., Gazaliyeva, M., Mustafin, K.,
Bedelbayeva, G., Dukenbayeva, B., Otarbayev, N. and Tordai, A., 2022.
Association between COVID-19 morbidity, mortality, and gross domestic
product, overweight/obesity, non-communicable diseases, vaccination
rate: A cross-sectional study. Journal of infection and public health,
15(2), pp.255-260.

Ren, M. (2020) Kaggle, COVID-19 Healthy Diet Dataset, kaggle.com,
accessed 17 October 2022.

Templin, T., Cravo Oliveira Hashiguchi, T., Thomson, B., Dieleman, J.
and Bendavid, E., 2019. The overweight and obesity transition from the
wealthy to the poor in low-and middle-income countries: A survey of
household data from 103 countries. PLoS medicine, 16(11), p.e1002968.
