---
title: "Project 603"
author: "Jannatul Naeema"
date: "2022-12-04"
output: html_document
---

```{r}

library(ggplot2)
library(lmtest)
library(mctest) 
library(GGally)
library(olsrr)
```



```{r}
dataset= read.csv("cleaned603dataset_-_dataset603_1.csv")
dataset
options(scipen = 999)
dataset <- na.omit(dataset)

```

Model:
```{r}
library(mctest)

fullmodel <-lm(Site_Energy~factor(Property_Type)+NumberofBuildings+Year_Built+Property_GFA+WeatherNormalizedSiteEnergyUse+SiteEUI+WeatherNormalizedSiteEUI+SourceEnergyUse+WeatherNormalizedSourceEnergyUse+SourceEUI+WeatherNormalizedSource.EUI+Emissions_CO2+Emissions_Intensity+DirectGHGEmissions+DirectGHGEmissionsIntensity+Electricity+Natural_Gas, data=dataset)

summary(fullmodel)
imcdiag(fullmodel, method='VIF')
```




```{r}

#fullmodel <-lm(Site_Energy~factor(Property_Type)+NumberofBuildings+Year_Built+Property_GFA+WeatherNormalizedSiteEnergyUse+SiteEUI+WeatherNormalizedSiteEUI+SourceEnergyUse+WeatherNormalizedSourceEnergyUse+SourceEUI+WeatherNormalizedSource.EUI+Emissions_CO2+Emissions_Intensity+DirectGHGEmissions+DirectGHGEmissionsIntensity+Electricity+Natural_Gas, data=dataset)

fullmodel <-lm(Site_Energy~NumberofBuildings+Property_GFA+Emissions_CO2+ Natural_Gas, data=dataset)
#fullmodel = lm(Site_Energy ~ NumberofBuildings+Year_Built+SourceEnergyUse+SourceEUI+Natural_Gas
#                  , data =energy)
summary(fullmodel)
imcdiag(fullmodel, method='VIF')
```
```{r}
fullmodel1 <-lm(Site_Energy~NumberofBuildings+Year_Built+SourceEnergyUse+SourceEUI+
                 Electricity, data=dataset)
#fullmodel = lm(Site_Energy ~ NumberofBuildings+Year_Built+SourceEnergyUse+SourceEUI+Natural_Gas
#                  , data =energy)
summary(fullmodel1)
imcdiag(fullmodel1, method='VIF')
```

```{r}
library(olsrr)

stepmod=ols_step_both_p(fullmodel,pent = 0.1, prem = 0.3, details=TRUE)
summary(stepmod$model)
```

Our fullmodel:
$$
\begin{aligned}
Y_{SiteEnergy}&=\beta_0 +  \beta_1 X_{SourceEnergyUse}  +\beta_2 X_{Electricity} +\beta_3 X_{SourceEUI}+\beta_4 X_{NumberofBuildings} +\epsilon  \\

\end{aligned}
$$
##Global F Test on Full Model:
Hypothesis Statement for Individual T-test:
$$
\begin{aligned}
H_0&:\beta_i=0\\
H_a&:\mbox{at least one }\beta_i\mbox{ is not zero } (i=1,2,3,4) 
\end{aligned}
$$

We set up the significance level at 0.05($\alpha = 0.05$).

Full Model Global F Test:
```{r}
#Full Model Test
fullmodel <-lm(Site_Energy~NumberofBuildings+Property_GFA+Emissions_CO2+ Natural_Gas, data=dataset)
reg<-lm(Site_Energy~1, data=dataset) # Model with only intercept

anova(fullmodel,reg) # We compare the NULL model with the full model
```
By using global F test, the output shows that $F_{cal}= 25794$ with df= -4, and $p-value< 0.00000000000000022 < \alpha=0.05$,indicating that we should clearly reject the null hypothesis. It provides compelling evidence against the null hypothesis. The Global F-test suggests that at least one of the independent variables must be related to Site Energy. Based on the p-value, we also have extremely strong evidence that at least one of the independent variables is associated with increased Site_Energy.

After we check the global F-test and reject the null hypothesis, we are checking the test statistics for the individual coefficients and particular subsets of the full model test in the following steps.

##Individual Coefficients Test:

###Hypothesis Statement:

$$
\begin{aligned}
H_0&:\beta_i=0\\
H_a&:\beta_i\neq0 (i=1,2,3,4)
\end{aligned}
$$
I set up the $\alpha = 0.05$

```{r}
summary(fullmodel)
```
From the out put, It shows that the variables
NumberofBuildings has $t_{cal}=-6.182$ with the $p-value = 0.00000000218 < 0.05$,
Property_GFA has $t_{cal}=11.761$ with the $p-value< 0.0000000000000002 < 0.05$  ,
Emissions_CO2 has $t_{cal}=58.698$ with the $p-value< 0.0000000000000002 < 0.05$,
Natural_Gas has $t_{cal}=122.447$ with the $p-value< 0.0000000000000002 < 0.05$ ,
indicating that we should clearly reject the null hypothesis for variable NumberofBuildings, Property_GFA,Emissions_CO2 and Natural_Gas, which means that we should keep those variables to the model at the significance level at $α=0.05$.


## Stepwise Regression Procedure

```{r}
#Stepwise Regression Procedure
library(olsrr)
library(readr)
library(dplyr)
backmodel=ols_step_backward_p(fullmodel, prem = 0.3)
summary(backmodel$model)

```

# Interaction Model
## Interaction Term T-tests:

```{r}
#interactions
#T-tests:
fullmodel <-lm(Site_Energy~NumberofBuildings+Property_GFA+Emissions_CO2+
                 Natural_Gas, data=dataset)
interactionmodel = lm(Site_Energy ~ (NumberofBuildings+Property_GFA+Emissions_CO2+
                 Natural_Gas)^2, data =dataset)
summary(interactionmodel)
```
```{r}
interactionmode2 = lm(Site_Energy ~ NumberofBuildings+Property_GFA+Emissions_CO2+
                 Natural_Gas+NumberofBuildings:Property_GFA+Property_GFA:Emissions_CO2+Property_GFA:Natural_Gas+Emissions_CO2:Natural_Gas, data =dataset)
summary(interactionmode2)
```


```{r}
interactionmode2 = lm(Site_Energy ~ NumberofBuildings+Property_GFA+Emissions_CO2+
                 Natural_Gas+Property_GFA:Emissions_CO2+Property_GFA:Natural_Gas+Emissions_CO2:Natural_Gas, data =dataset)
summary(interactionmode2)
```



```{r}
#Second Order Term T-tests:

qmodel = lm(Site_Energy ~ Number.of.Buildings+I(Number.of.Buildings^2)+Source.Energy.Use..GJ.+I(Source.Energy.Use..GJ.^2)+Source.EUI..GJ.m..+I(Source.EUI..GJ.m..^2)+Electricity+I(Electricity^2)
                  , data =energy)

summary(qmodel)
```
```{r}
qmodel2 = lm(Site_Energy ~ Number.of.Buildings+Source.Energy.Use..GJ.+I(Source.Energy.Use..GJ.^2)+I(Source.Energy.Use..GJ.^3)+I(Source.Energy.Use..GJ.^4)+Source.EUI..GJ.m..+Electricity+I(Electricity^2)+I(Electricity^3)+I(Electricity^4)+I(Electricity^5)+I(Electricity^6)+I(Electricity^7)
                  , data =energy)

summary(qmodel2)

```






```{r}
bestm = lm(Site_Energy ~ Number.of.Buildings+Source.Energy.Use..GJ.+I(Source.Energy.Use..GJ.^2)+I(Source.Energy.Use..GJ.^3)+I(Source.Energy.Use..GJ.^4)+Source.EUI..GJ.m..+Electricity+I(Electricity^2)+I(Electricity^3)+I(Electricity^4)+I(Electricity^5)+I(Electricity^6)+I(Electricity^7)+ Number.of.Buildings:Source.Energy.Use..GJ. + Source.Energy.Use..GJ.:Electricity + Source.EUI..GJ.m..:Electricity
                  , data =energy)
summary(bestm)
```
```{r}
bestm = lm(Site_Energy ~ Number.of.Buildings+Source.Energy.Use..GJ.+I(Source.Energy.Use..GJ.^2)+I(Source.Energy.Use..GJ.^3)+Source.EUI..GJ.m..+Electricity+I(Electricity^2)+ Number.of.Buildings:Source.Energy.Use..GJ. + Source.EUI..GJ.m..:Electricity
                  , data =energy)
summary(bestm)
```
```{r}
#Higher Order Partial F-tests:
anova(reducedinteractionmodel2, bestm)
```
```{r}
#Linearity Assumption

library(ggplot2)
ggplot(bestm, aes(x=.fitted, y=.resid)) +
  geom_point() + geom_smooth()+
  geom_hline(yintercept = 0)+
  ggtitle("Residual vs Fitted")
```
```{r}
#Normality Assumption
shapiro.test(residuals(bestm))
qplot(residuals(bestm),
      geom="histogram",
      binwidth = 100,  
      main = "Histogram of residuals", 
      xlab = "residuals", color="red", 
      fill=I("blue"))

ggplot(energy, aes(sample=bestm$residuals)) +
  stat_qq() +
  stat_qq_line() +
  ggtitle("Normal Q-Q Plot")
```
```{r}
#Equal Variance Assumption
library(lmtest)
bptest(bestm)
plot(bestm,which=1)

```
```{r}
#Multicollinearity Assumption
library(mctest)
library(ggplot2)
library(GGally)
fullmodel = lm(Site_Energy ~ Number.of.Buildings+Year_Built+Source.Energy.Use..GJ.+Source.EUI..GJ.m..+Electricity
                  , data =energy)#ummary(fullcar2)
catfp2 <- data.frame(energy$Number.of.Buildings,energy$Year_Built,energy$Source.Energy.Use..GJ.,energy$Source.EUI..GJ.m..,energy$Electricity)
ggpairs(catfp2,lower = list(continuous = "smooth_loess", combo = "facethist", discrete = "facetbar", na = "na"),upper = list(continuous = wrap("cor", family="sans")))
```
```{r}
#Influential Outliers

plot(bestm,which=5)
```
```{r}
car2[cooks.distance(bestm)>0.5,]
plot(bestm,pch=18,col="red",which=c(4))

```


